---
title: "Crime Types from a Gender Perspective Forced Marriage & Female Genital Mutilation"
author: "Eunkoo Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  word_document: default
  pdf_document: default
  html_document:
    code_folding: hide
    highlight: tango
    number_sections: yes
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set( message = FALSE, warning = FALSE, results =FALSE, echo = TRUE) 

```
**Important notice:**

In this document the top right button **CODE**  enables you to show all R code chunks that have been hidden by default. You can choose to show the code either individually or document wide using these buttons.

----

*R codes for generating child marriage/FGM indicators in the Sudan MICS 2014 data can be found on the [SIAP e-learning Platform](https://siap-elearning.org/index.php?)*

R code is available [here](FGMsudan.R) 

---

# Installation

We use some standard packages for doing the analysis. These packages have to be downloaded and installed before loading them here. (see **code** ) 


```{r packages, echo = TRUE}
# loading libraries. Might need to install packages first if not installed already

library(foreign)
library(plyr)
library(dplyr)
library(tidyr)
library(tidyverse)
library(ggplot2)

```

# Reading data


```{r data}
#-------------------
# See with Eunkoo how to settle this since Rmarkdown has a different logic. (path are relative to the folder where the .Rmd is stored) 
#-------------------
# setting the working directory by specifying the path of the folder ## To discuss with Eunkoo 
# setwd("C:/Users/exl15/Downloads/Rwork/sudan")
#setwd("C:/Chris/Un")

# checking the path of current directory
#getwd()

# reading MICS micro-data into R
# reading micro-data from questionnaire for women and storing into wm
wm<-read.spss(file="Data/wm.sav",to.data.frame = TRUE)

```



# Initial Analysis 
We can perform some analysis from the data set. 
First, we can check that after reading the data, there are **`r nrow(wm)`** observations in this file, and **`r ncol(wm)`** variables. 

## Child marriage
We can compute some initial counts and percentages: (see **code** ) 

```{r inital, results= FALSE}
#subset function takes parts of dataset from “wm" dataset that meets certain #conditions. The conditions are women who have completed the questionnaire #(WM7=Completed) and their age is 20-24 which is the denominator. #Numerator includes number of women first married or in union before age 18 #(WAGEM). 

wmcompleted<-subset(wm, (WM7=='Completed') & WB2>=20 & WB2<25) 

sum(wmcompleted$wmweight)

age18<-subset(wmcompleted, WAGEM<18)

sum(age18$wmweight)/sum(wmcompleted$wmweight)

```

We take a subset from the original file. The conditions are women who have completed the questionnaire (*WM7=Completed*) and their age is *20-24* which is the denominator. 

The total (weighted)number of completed cases is  **`r round(sum(wmcompleted$wmweight),2)`** .

We can compute a percentage using as numerator the number of women first married or in union before age 18 (WAGEM). 

> Within those cases, the percentage of women under 18  is  **`r round(100*sum(age18$wmweight)/sum(wmcompleted$wmweight), 2)`** \%. 


## Percent currently married (age 15-19)

Using the same type of code, we can compute the percentage of married or formerly married women. (see **code** )
```{r}
# Numerator includes women aged 15-19 who completed the questionnaire and the 
# marital status (MSTATUS) is “currently married” or “formerly married”. Denominator 
# includes all women aged 15-19 who completed the questionnaire. 
# In R, logical operator “|” is used for OR condition. 
# For AND condition, logical operator “&” is used.

wmcompleted<-subset(wm, (WM7=='Completed') & WB2>=15 & WB2<20)

cmarried<-subset(wmcompleted, (MSTATUS=='Currently married')|(MSTATUS=='Formerly married'))

sum(cmarried$wmweight)/sum(wmcompleted$wmweight)

```
We can now say that:

> The percentage of women currently or formerly married is **`r round(100*sum(cmarried$wmweight)/sum(wmcompleted$wmweight),2)`** %


*Visualization proposition* 

```{r plotMarried}
wmCB <- wm %>% 
  filter(WM7=='Completed' & WB2>=15 & WB2<20) %>%
  group_by(MSTATUS) %>% 
  summarise(count = n()) %>% 
  mutate(perc = 100*count/sum(count)) %>%
  ungroup()

ggplot(wmCB, aes(x = factor(MSTATUS), y = perc)) +
  geom_bar(stat="identity", width = 0.7, fill = "steelblue") +
labs(title = "Marital Status of women (age 15-19)",
        x =  "Marital Status",
        y = "percent") +
  theme_minimal()

```

## Percent currently married (age 15-19): education

In the code used,  Numerator includes women aged 15-19 who completed the questionnaire and the  marital status (MSTATUS) is “currently married” or “formerly married”. Denominator   includes all women aged 15-19 who completed the questionnaire. 

In R, logical operator “|” is used for OR condition. For AND condition, logical operator “&” is used. (see **code**)



```{r}
wmcompleted<-subset(wm, (WM7=='Completed') & WB2>=15 & WB2<20)
edunone<-subset(wmcompleted, welevel=='None')
eduprimary<-subset(wmcompleted, welevel=='Primary')
edusecondary<-subset(wmcompleted, welevel=='Secondary')
eduhigher<-subset(wmcompleted, welevel=='Higher')

cmedunone<-subset(edunone, (MSTATUS=='Currently married')|(MSTATUS=='Formerly married'))
cmeduprimary<-subset(eduprimary, (MSTATUS=='Currently married')|(MSTATUS=='Formerly married'))
cmedusecondary<-subset(edusecondary, (MSTATUS=='Currently married')|(MSTATUS=='Formerly married'))
cmeduhigher<-subset(eduhigher, (MSTATUS=='Currently married')|(MSTATUS=='Formerly married'))


sum(cmedunone$wmweight)/sum(edunone$wmweight) #FGM prevalence ratio none
sum(cmeduprimary$wmweight)/sum(eduprimary$wmweight) #FGM prevalence ratio primary
sum(cmedusecondary$wmweight)/sum(edusecondary$wmweight) #FGM prevalence ratio secondary
sum(cmeduhigher$wmweight)/sum(eduhigher$wmweight) #FGM prevalence ratio hihger

```
 **Bla bla... **
 
* FGM prevalence ratio none:
* FGM prevalence ratio primary:
* FGM prevalence ratio:
* FGM prevalence ratio higher:

*Visualisation proposition* 

```{r marriage}
ggplot(wmcompleted) +
 aes(x = welevel, fill = MSTATUS) +
 geom_bar() +
 scale_fill_hue() +
 labs(x = "Education", title = "Marital Status and education", 
      fill = "Marital Status") +
 theme_minimal()
```

## Percent currently married (age 15-19): urban/rural

```{r}
wmcompleted<-subset(wm, (WM7=='Completed') & WB2>=15 & WB2<20)
wmcompletedu<-subset(wm, (WM7=='Completed') & (WB2>=15 & WB2<20)&(HH6=='Urban'))
wmcompletedr<-subset(wm, (WM7=='Completed') & (WB2>=15 & WB2<20)&(HH6=='Rural'))
Cmarriedu<-subset(wmcompletedu, (MSTATUS=='Currently married')|(MSTATUS=='Formerly married'))
cmarriedr<-subset(wmcompletedr, (MSTATUS=='Currently married')|(MSTATUS=='Formerly married'))
Cmarriedu<-subset(cmarried, HH6=='Urban')
Cmarriedr<-subset(cmarried, HH6=='Rural')
sum(Cmarriedu$wmweight)/sum(wmcompletedu$wmweight)
sum(Cmarriedr$wmweight)/sum(wmcompletedr$wmweight)
```

*A proposal for visualizing some numbers*

```{r, plotLiving, result = TRUE}
wmCB<- wm %>%
  filter(WM7=='Completed' & WB2>=15 & WB2<20) %>%
  mutate(LivingArea = ifelse(HH6=='Urban',"URBAN", ifelse(HH6=='Rural', "RURAL", NA_character_) )
  )
   ggplot(wmCB, aes(LivingArea)) +
     geom_bar(fill="steelblue")+
     theme_minimal()
   
   
   
      
```





### SDG 5.3.2:FGM

```{r}
wmcompleted<-subset(wm, WM7=='Completed')
sum(wmcompleted$wmweight)
wmyes<-subset(wmcompleted, FG3=='Yes')
sum(wmyes$wmweight)
sum(wmyes$wmweight)/sum(wmcompleted$wmweight) #FGM prevalence ratio

```


## FGM by urban/rural
```{r}
urbanity<-subset(wmcompleted, HH6=='Urban') #denominator for urban
rurality<-subset(wmcompleted, HH6=='Rural') #denominator for rural
sum(urbanity$wmweight)
sum(rurality$wmweight)
urbanityyes<-subset(urbanity, FG3=='Yes') #numerator for urban
ruralityyes<-subset(rurality, FG3=='Yes') #numerator for rural
sum(urbanityyes$wmweight)/sum(urbanity$wmweight) #FGM prevalence ratio urban
sum(ruralityyes$wmweight)/sum(rurality$wmweight) #FGM prevalence ratio rural

```

**TBC...** 
